Uncovering underlying high-level musical content in the time domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn deep audio embeddings with applications to MIR downstream tasks

### 2023-06-11 evaluation/evaluation.tex
Updates from Overleaf

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Evaluation dataset and metrics}

The developed model has been evaluated on the SALAMI dataset \cite{Smith2011DESIGNANNOTATIONS}. The "Structural Analysis of Large Amounts of Music Information" project aims to execute comprehensive structural analyses on a diverse spectrum of music. In contrast to traditional methods, SALAMI segments musical pieces into distinct sections, integrating perceptual, functional, and transcription analyses. Despite certain limitations, this novel approach offers a comprehensive and nuanced understanding of musical structure.

Three primary metrics have been utilized to assess the model performance: Precision, Recall, and F-measure. Precision quantifies the proportion of accurately identified boundaries relative to all estimated boundaries to indicate the algorithm's accuracy in boundary detection.

\begin{equation}
\text{Precision} = \frac{\text{True Positives}}{\text{True Positives} + \text{False Positives}}
\end{equation}

On the other hand, Recall measures the proportion of accurately detected boundaries against all reference boundaries, indicating the completeness or sensitivity of the algorithm.

\begin{equation}
\text{Recall} = \frac{\text{True Positives}}{\text{True Positives} + \text{False Negatives}}
\end{equation}

Lastly, the F-measure offers a harmonized view of precision and recall. A widely adopted boundary detection metric, it compares predicted and ground truth boundaries, outputting a score between 0 and 1. This score is calculated as the harmonic mean of Precision and Recall, effectively considering both under-segmentation and over-segmentation. Given the potential inaccuracies inherent in human annotations and prediction errors, the F-measure permits minor deviations between predicted and actual boundaries. Its tolerance threshold can be adjusted, allowing a predicted boundary to be deemed correct if it is within a predefined window of a ground truth boundary \cite{Turnbull2007ABOOSTING, NietoPerceptualMusicb}.

\begin{equation}
\text{F-measure} = \frac{2 \times \text{Precision} \times \text{Recall}}{\text{Precision} + \text{Recall}}
\end{equation}

### 2023-06-11 figures/neural networks/colour CNN.tex
Updates from Overleaf

\begin{figure}[ht]
    \centering
% DEEP CONVOLUTIONAL NEURAL NETWORK
\begin{tikzpicture}[x=1.6cm,y=1.1cm]
  \large
  \message{^^JDeep convolution neural network}
  \readlist\Nnod{5,5,4,3,2,4,4,3} % array of number of nodes per layer
  \def\NC{6} % number of convolutional layers
  \def\nstyle{int(\lay<\Nnodlen?(\lay<\NC?min(2,\lay):3):4)} % map layer number on 1, 2, or 3
  \tikzset{ % node styles, numbered for easy mapping with \nstyle
    node 1/.style={node in},
    node 2/.style={node convol},
    node 3/.style={node hidden},
    node 4/.style={node out},
  }

@@

\end{tikzpicture}
    \caption{High-level illustration of a CNN}
    \label{fig: CNN colour diagram}
\end{figure}

>>>>

\begin{figure}[ht]
    \centering
% DEEP CONVOLUTIONAL NEURAL NETWORK
\begin{tikzpicture}[x=1.5cm,y=1.0cm]
  \large
  \message{^^JDeep convolution neural network}
  \readlist\Nnod{5,5,4,3,2,4,4,3} % array of number of nodes per layer
  \def\NC{6} % number of convolutional layers
  \def\nstyle{int(\lay<\Nnodlen?(\lay<\NC?min(2,\lay):3):4)} % map layer number on 1, 2, or 3
  \tikzset{ % node styles, numbered for easy mapping with \nstyle
    node 1/.style={node in},
    node 2/.style={node convol},
    node 3/.style={node hidden},
    node 4/.style={node out},
  }

@@

\end{tikzpicture}
    \caption[High-level CNN diagram \cite{tikz}.]{\small{High-level illustration of a CNN. It consists of an input layer that takes in numerical data, followed by a downsampling process using convolution and pooling layers to extract features. Finally, fully-connected hidden layers are added for learning the global structure, and an output layer produces the desired predictions.}}
    \label{fig: CNN colour diagram}
\end{figure}

### 2023-06-11 future work/future work.tex
Updates from Overleaf

The question as per \cite{Turian2022HEAR:Representations} or \cite{Li2023MERT:Training} remains if such general-purpose audio representation can mimic human hearing.

transforms
stems
different takes
k-fold cross validation
hyper-params: kernel size 0.005 times the sample rate

Utilize the dB-scale Mel-spectrum magnitude of audio as input data; it is a popular choice for input representation in music-related tasks when applying CNNs, as has been demonstrated in various studies \cite{Kim2020OneStrategies}. In addition to their wide usage, it's been recently reported that their frequency-domain summarization, grounded in psycho-acoustics, is efficient and challenging to replicate through solely data-driven methods. \cite{Kim2020OneStrategies}

>>>>

The question as per \cite{Turian2022HEAR:Representations} or \cite{Li2023MERT:Training} remains if such general-purpose audio representation can mimic human hearing. That is a "thorn in our side" that we will pursue.

transforms
stems
different takes
k-fold cross validation
hyper-params: kernel size 0.005 times the sample rate
larger representation layer
Loss function margin. A margin of 0.2 can be a starting point, but it might not be the best value for every task. It's recommended to experiment with different values and choose the one that provides the best performance on your validation set.

Utilize the dB-scale Mel-spectrum magnitude of audio as input data; it is a popular choice for input representation in music-related tasks when applying CNNs, as has been demonstrated in various studies \cite{Kim2020OneStrategies}. While raw audio is the most meaningful audio representation, it's been reported that dB-scale Mel-spectrum frequency-domain summarization, grounded in psycho-acoustics, is computationally efficient and challenging to replicate solely through data-driven methods \cite{Kim2020OneStrategies}; therefore, a trade-off that looks worthy of being explored.

### 2023-06-11 introduction/Subsections/Previous work.tex
Updates from Overleaf

Given their demonstrated effectiveness in existing literature and their potential for transfer learning, deep audio embeddings hold substantial promise for MIR downstream tasks, specifically boundary detection. Nevertheless, it remains uncertain whether a general-purpose audio representation can effectively replicate human hearing \cite{Turian2022HEAR:Representations}, even though some techniques have managed to generalize across up to fourteen music understanding tasks \cite{Li2023MERT:Training}.

>>>>

Given their demonstrated effectiveness in the existing literature and their potential for transfer learning, deep audio embeddings hold substantial promise for MIR downstream tasks, and boundary detection is no exception. Nevertheless, it remains uncertain whether a general-purpose audio representation can effectively replicate human hearing \cite{Turian2022HEAR:Representations}, even though some techniques have managed to generalize across up to fourteen music understanding tasks \cite{Li2023MERT:Training}.

### 2023-06-11 methods/subsections/deep learning.tex
Updates from Overleaf

\section{Deep Learning Modeling}

@@

Deep Neural Networks (DNNs), computational models rooted in biological neural systems, utilize layers of artificial neurons for tasks such as pattern recognition, classification, and regression. DNNs evolve by adjusting connections between neurons during learning, a concept initiated in the mid-20th century but modernized with Frank Rosenblatt's single-layer perceptron in 1957, capable of learning linearly separable patterns.

>>>>

\section{Deep Learning modeling}

@@

Deep Neural Networks (DNNs), computational models somewhat rooted in biological neural systems, utilize layers of artificial neurons for tasks such as pattern recognition, classification, and regression. DNNs evolve by adjusting connections between neurons during learning, a concept initiated in the mid-20th century but modernized with Frank Rosenblatt's single-layer perceptron in 1957, capable of learning linearly separable patterns \cite{perceptron}.

### 2023-06-11 methods/subsections/implementation.tex
Updates from Overleaf

\subsection{Triplet Siamese Networks}

This work suggests employing a Triple Siamese Network, a model architecture known for its efficacy in music similarity retrieval tasks \cite{contentmusicsimtriplet2020}. The aim is to minimize the loss function between a triplet of anchor, positive, and negative samples, utilizing online triplet mining for optimizing memory resources \cite{Sikaroudi2020OfflinePatches}. This self-supervised learning (SSL) model aims to distinguish between similar and dissimilar samples effectively.

@@

@@

The triplet loss function used in Siamese-like networks guides the learning process. It seeks to minimize the distance between the anchor and positive instances while maximizing the distance between the anchor and the negative instances. A margin parameter is included in the loss function to ensure a minimum separation between positive and negative instances in the embedding space.

The triplet loss function $\mathcal{L}(\mathbf{a}, \mathbf{p}, \mathbf{n})$ aims to ensure that an anchor vector $\mathbf{a}_i$ is closer in the embedding space to a positive vector $\mathbf{p}_i$ (representing an example of the same class) than to a negative vector $\mathbf{n}_i$ (representing an example of a different class) by at least a margin $m$. It is calculated by summing the losses overall $N$ triplets in the dataset, where the equation gives the loss for each triplet:

\begin{equation}
\mathcal{L}(\mathbf{a}, \mathbf{p}, \mathbf{n}) = \sum_{i=1}^{N} \max \left(0, \left| \mathbf{a}_i - \mathbf{p}_i \right|_2^2 - \left| \mathbf{a}_i - \mathbf{n}_i \right|_2^2 + m \right)
\end{equation}

The final loss used for model training is then the average loss over a mini-batch of $N$ triplets:

@@

@@

$I$ represents the set of all items in the batch, $|A_i|$, $|P_i|$, and $|N_i|$ denote the lengths of the anchor, positive, and negative embedding vectors for the $i$-th item, respectively. The $\max$ function is applied to find the longest of these three lengths for each item, and then the maximum of these maximum lengths is taken over all items in the batch. This gives the maximum length, $L_{\text{max}}$, of any vector in the batch.

>>>>

\subsection{Deep Neural Network Architecture}

This work suggests employing a Triple Siamese Network, a model architecture known for its efficacy in music similarity retrieval tasks \cite{contentmusicsimtriplet2020}. The aim is to minimize the loss function between a triplet of anchor, positive, and negative samples, utilizing online triplet mining for optimizing memory resources \cite{Sikaroudi2020OfflinePatches}. This SSL model aims to distinguish between similar and dissimilar samples effectively.

@@

There are several loss functions used in tasks like feature learning, similarity learning, and dimensionality reduction.

Functions such Mean Squared Error (MSE) Loss \cite{Hodson2021MeanDeconstructed}, Cross-Entropy Loss \cite{Zhang2018GeneralizedLabels}, Contrastive Loss \cite{supercontrast} or Triplet Loss are used when training models to understand the difference between various classes or categories.

@@

The triplet loss function guides the learning process. It seeks to minimize the distance between the anchor and positive instances while maximizing the distance between the anchor and the negative instances. A margin parameter is included in the loss function to ensure a minimum separation between positive and negative instances in the embedding space.

The triplet loss function $\mathcal{L}(\mathbf{a}, \mathbf{p}, \mathbf{n})$ aims to ensure that an anchor vector $\mathbf{a}_i$ is closer in the embedding space to a positive vector $\mathbf{p}_i$ (representing an example of the same class) than to a negative vector $\mathbf{n}_i$ (representing an example of a different class) by at least a margin $\alpha$. It is calculated by summing the losses overall $N$ triplets in the dataset, where the equation gives the loss for each triplet:

\begin{equation}
\mathcal{L}(\mathbf{a}, \mathbf{p}, \mathbf{n}) = \sum_{i=1}^{N} \max \left(0, \left| \mathbf{a}_i - \mathbf{p}_i \right|_2^2 - \left| \mathbf{a}_i - \mathbf{n}_i \right|_2^2 + \alpha \right)
\end{equation}

The $\max(0, x)$ operation ensures zero loss when the distances satisfy this condition. The final loss used for model training is then the average loss over a mini-batch of $N$ triplets:

@@

The margin is a task-dependent optimal value determined empirically based on model performance. If it's too small, the model might not differentiate classes effectively; if it's too large, it might focus on outliers.

@@

$I$ represents the set of all items in the batch, $|A_i|$, $|P_i|$, and $|N_i|$ denote the lengths of the anchor, positive, and negative vectors for the $i$-th item, respectively. The $\max$ function is applied to find the longest of these three lengths for each item, and then the maximum of these maximum lengths is taken over all items in the batch. This gives the maximum length, $L_{\text{max}}$, of any vector in the batch.

\subsubsection{Training strategy}

To tackle the computational demands of training models on extensive raw audio data, we incorporated a couple of strategies to optimize efficiency and performance.

First, we utilized 16-mixed precision training. This approach leverages the improved performance of modern GPUs for 16-bit computations, enabling the model to run faster and use less memory without sacrificing model performance \cite{Das2018MixedOperations}.

Secondly, to capitalize on the computational capabilities of multiple GPUs and hasten training times, we employed the Distributed Data-Parallel (DDP) strategy \cite{Li2020PyTorchTraining}. DDP operates on distinct mini-batches of data across GPUs and synchronizes the gradients after each backward pass, providing a more efficient scaling than other parallel strategies.

These strategies collectively enhance the computational efficiency while maintaining the robustness of the model training on lengthy raw audio data.

### 2023-06-11 results/subsections/tables.tex
Updates from Overleaf

SALAMI (Structural Analysis of Large Amounts of Music Information) is a project aiming to carry out extensive structural analyses on a wide variety of music, from Western pop to Indian classical, and from live to studio recordings \cite{Smith2011DESIGNANNOTATIONS}. Instead of traditional forms, the authors segment pieces into sections, combining perceptual, functional, and transcription analyses. Despite limitations, this approach offers a holistic and nuanced perspective on music structure.

@@

\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{|X|X|X|X|X|}
\hline
\textbf{Technique} & \textbf{Training Dataset} & \textbf{Precision} & \textbf{Recall} & \textbf{F-measure} \\
\hline
MFCC & -- & -- & -- & -- \\
\hline
Embeddiogram & GTZAN Dataset & -- & -- & -- \\
\hline
Embeddiogram & MSD Dataset & -- & -- & -- \\
\hline
PCP & -- & -- & -- & -- \\
\hline
Tonnetx & -- & -- & -- & -- \\
\hline
CQT & -- & -- & -- & -- \\
\hline
\end{tabularx}
\caption{Comparison of precision, recall, and f-measure for different techniques using the SALAMI dataset and the training dataset used for each technique.}
\label{tab:comparison}
\end{table}

@@

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{4.5cm}XXXXX}
\toprule
\thead{\centering\textbf{Authors [Ref], Year}} & \thead{\centering\textbf{Input}} & \thead{\centering\textbf{Method}} & \thead{\centering\textbf{F-measure}} \\
\midrule
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & Constant-Q Transform Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
Grill \& Schlüter \cite{GrillMUSICANNOTATIONS}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State-of-the-art table.]{\small{Modified table presented by Hernandez-Olivan, C., Beltran, J. R. \& Diaz-Guerra, D. in \cite{Hernandez-Olivan2021MusicFeatures}. Previous studies' boundary detection f-measure results using unsupervised methods for a 0.5s time-window tolerance. Only the top-performing algorithm for each year on the SALAMI dataset is displayed.}}
\label{tab:my_label}
\end{table}

>>>>

SALAMI (Structural Analysis of Large Amounts of Music Information) is a project aiming to carry out extensive structural analyses on a wide variety of music \cite{Smith2011DESIGNANNOTATIONS}. Instead of traditional forms, the authors segment pieces into sections, combining perceptual, functional, and transcription analyses. Despite limitations, this approach offers a holistic and nuanced perspective on music structure.

@@

\begin{table}[h]
\centering
\begin{tabularx}{\textwidth}{|X|X|X|X|X|}
\hline
\textbf{Technique} & \textbf{Training Dataset} & \textbf{Precision} & \textbf{Recall} & \textbf{F-measure} \\
\hline
MFCC & -- & -- & -- & -- \\
\hline
Embeddiogram & GTZAN Dataset & -- & -- & -- \\
\hline
Embeddiogram & MSD Dataset & -- & -- & -- \\
\hline
PCP & -- & -- & -- & -- \\
\hline
Tonnetz & -- & -- & -- & -- \\
\hline
CQT & -- & -- & -- & -- \\
\hline
\end{tabularx}
\caption{Comparison of precision, recall, and f-measure for different techniques using the SALAMI dataset and the training dataset used for each technique.}
\label{tab:comparison}
\end{table}

@@

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{4.5cm}XXXXX}
\toprule
\thead{\centering\textbf{Authors [Ref], Year}} & \thead{\centering\textbf{Input}} & \thead{\centering\textbf{Method}} & \thead{\centering\textbf{F-measure}} \\
\midrule
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & Constant-Q Transform Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
Grill \& Schlüter \cite{GrillMUSICANNOTATIONS}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State-of-the-art table.]{\small{Previous studies' boundary detection f-measure results using unsupervised methods for a 0.5s time-window tolerance. Only the top-performing algorithm for each year on the SALAMI dataset is displayed. Original source \cite{Hernandez-Olivan2021MusicFeatures}}}
\label{tab:my_label}
\end{table}
